---
title: "Creation of Series Averages DataFrames"
output: html_notebook
---

The csv file [championsdata.csv](http://daverosenman.com/championsdata.csv) and [runnerupsdata.csv](http://daverosenman.com/runnerupsdata.csv) contain team totals from every NBA finals game since 1980. 'championsdata.csv' contains the team totals for the championship teams and 'runnerupsdata.csv' contains the teams totals for the losing team. 
I wrangled those files from basketball-reference.com. 

From the championsdata.csv and runnerupsdata.csv datasets, I created the following datasets:
- [champs_and_runner_ups_series_averages.csv](http://daverosenman.com/champs_and_runner_ups_series_averages.csv): Data set containing series averages for both the champs and runner-up.
- [champs_series_averages.csv](http://daverosenman.com/champs_series_averages.csv): Data set containing series averages for the champs only.
- [runner_ups_series_averages.csv](http://daverosenman.com/runner_ups_series_averages.csv): Data set containing series averages for the runner-ups only.

In the cells below, I created champs_and_runner_ups_series_averages.csv, champs_series_averages.csv, and runner_ups_series_averages.csv.

```{r}
suppressMessages(library(tidyverse))
library(readr)
```

```{r}
#NBA Finals game-by-game data
#Champions Data
champs <- readr::read_csv('championsdata.csv',col_types = cols())   
#Runner-ups Data
runner_ups <- read.csv('runnerupsdata.csv')
```

```{r}
#Creation of Single Column data.frame with champion names for later use
champ_team <- champs %>% filter(Game == 3) %>% select(Team)

#Creation of Single Column data.frme with runner-up names for later use
runner_up_team <- runner_ups %>% filter(Game == 3) %>% select(Team)

#Blending together champs and runner_ups dataframes
champs <- select(champs,-c(X,Team,FGP,TPP,FTP))
runner_ups <- select(runner_ups,-c(Y,Team,FGP,TPP,FTP))
champs_and_runner_ups_by_year <- data.frame(1980:2017,champ_team[1],runner_up_team[1])
colnames(champs_and_runner_ups_by_year) <-  c('Year','Champion','Runner-Up')
```

```{r}
#creation of dataframe with average values for champions and runner-ups in each finals series since 1980
champs_averages <- champs %>% group_by(Year) %>% summarize_each(funs(mean))
champs_averages['Status']  <-  'Champion'
champs_averages['Team']  <- champ_team
runner_ups_averages <- runner_ups %>% group_by(Year) %>% 
summarize_each(funs(mean))
runner_ups_averages['Status']  <-  'Runner Up'            
runner_ups_averages['Team']  <- runner_up_team
averages = rbind(champs_averages,runner_ups_averages)
averages  <- averages %>% arrange(Year,Status)

#dataframe of averages of champions and runner ups from each nba finals series since 1980
averages <- averages %>% select(Year,Status,Team,everything()) %>%select(-c(Game,Win,Home,MP))
```

```{r}
champs_and_runner_ups_series_averages <- averages %>% mutate(FGP = 100*FG/FGA,FTP = 100*FT/FTA,TPP = 100*TP/TPA)
champs_and_runner_ups_series_averages <- champs_and_runner_ups_series_averages %>% select(Year:Team,PTS,FG,FGA,FGP,TP,TPA,TPP,FT:FTA,FTP,everything())
champs_and_runner_ups_series_averages
```

```{r}
#Creation of csvs
write.csv(champs_and_runner_ups_series_averages,'champs_and_runner_ups_series_averages.csv')
write.csv(filter(champs_and_runner_ups_series_averages,Status == 'Champion'),'champs_series_averages.csv')
write.csv(filter(champs_and_runner_ups_series_averages,Status == 'Runner Up'),'runner_ups_series_averages.csv')
```


```{r}
#Creation of Averages Grouped By Decade

averages <- readr::read_csv('champs_and_runner_ups_series_averages.csv',col_types = cols())


#Average Three Pointers Made Per Game, Decade  
averages_over_years_TP <- function(start,end) {
  year_string <- paste(as.character(start),as.character(end),sep = '-')
  df <-  averages %>% filter((Year >= start) & (Year <= end)) %>% group_by(Status) %>% summarize(Average_Made_Threes = mean(TP)) %>% mutate(Years = year_string) %>% select(Years,Status, Average_Made_Threes)
  df 
}

i <-  0
tp_averages_by_decade = data.frame()
while (i < 3) {
  start <-  1980+10*i
  end <- 1989+10*i
  tp_averages_by_decade <- rbind(tp_averages_by_decade,averages_over_years_TP(start,end))
  i = i + 1
}
tp_averages_by_decade <- rbind(tp_averages_by_decade,averages_over_years_TP(2010,2017))
tp_averages_by_decade

```
```{r}
#Average Three Point Attempts Per Game, Decade
averages_over_years_TPA <- function(start,end) {
  year_string <- paste(as.character(start),as.character(end),sep = '-')
  df <-  averages %>% filter((Year >= start) & (Year <= end)) %>% group_by(Status) %>% summarize(Average_Attempted_Threes = mean(TPA)) %>% mutate(Years = year_string) %>% select(Years,Status, Average_Attempted_Threes)
  df 
}

i <-  0
tpa_averages_by_decade = data.frame()
while (i < 3) {
  start <-  1980+10*i
  end <- 1989+10*i
  tpa_averages_by_decade <- rbind(tpa_averages_by_decade,averages_over_years_TPA(start,end))
  i = i + 1
}
tpa_averages_by_decade <- rbind(tpa_averages_by_decade,averages_over_years_TPA(2010,2017))

tpa_averages_by_decade
```

```{r}
#Three Point Percentage, Grouped By Decade
years <- tpa_averages_by_decade$Years
status <- tpa_averages_by_decade$Status
tpp_by_decade <- data.frame(Years = years, Status = status,Three_Point_Percentage = 100*tp_averages_by_decade$Average_Made_Threes/tpa_averages_by_decade$Average_Attempted_Threes)
tpp_by_decade
```
```{r}
#Average Field Goals Made By Decade
averages_over_years_FG <- function(start,end) {
  year_string <- paste(as.character(start),as.character(end),sep = '-')
  df <-  averages %>% filter((Year >= start) & (Year <= end)) %>% group_by(Status) %>% summarize(Average_Made_Field_Goals = mean(FG)) %>% mutate(Years = year_string) %>% select(Years,Status, Average_Made_Field_Goals)
  df 
}

i <-  0
fg_averages_by_decade = data.frame()
while (i < 3) {
  start <-  1980+10*i
  end <- 1989+10*i
  fg_averages_by_decade <- rbind(fg_averages_by_decade,averages_over_years_FG(start,end))
  i = i + 1
}
fg_averages_by_decade <- rbind(fg_averages_by_decade,averages_over_years_FG(2010,2017))
fg_averages_by_decade
```

```{r}
#Average Field Goal Attempts by Decade
averages_over_years_FGA <- function(start,end) {
  year_string <- paste(as.character(start),as.character(end),sep = '-')
  df <-  averages %>% filter((Year >= start) & (Year <= end)) %>% group_by(Status) %>% summarize(Average_Attempted_Field_Goals = mean(FGA)) %>% mutate(Years = year_string) %>% select(Years,Status, Average_Attempted_Field_Goals)
  df 
}

i <-  0
fga_averages_by_decade = data.frame()
while (i < 3) {
  start <-  1980+10*i
  end <- 1989+10*i
  fga_averages_by_decade <- rbind(fga_averages_by_decade,averages_over_years_FGA(start,end))
  i = i + 1
}
fga_averages_by_decade <- rbind(fga_averages_by_decade,averages_over_years_FGA(2010,2017))
fga_averages_by_decade

```

```{r}
#Field Goal Percentage By Decade
fgp_by_decade <- data.frame(Years = years, Status = status, FGP = fg_averages_by_decade$Average_Made_Field_Goals/fga_averages_by_decade$Average_Attempted_Field_Goals*100)
fgp_by_decade
```


```{r}
#Percentage of Made Field Goals That Were 3's, By Decade


percentage_of_fgs_tps <- data.frame(Years = years,Status = status,Percentage_of_FGs_TPs = tp_averages_by_decade$Average_Made_Threes/fg_averages_by_decade$Average_Made_Field_Goals*100)
percentage_of_fgs_tps

```
```{r}
#Percentage of Field Goals Attempts That Were 3's, By Decade
percentage_of_fgas_tps <- data.frame(Years = years,Status = status,Percentage_of_FGAs_TPs = tpa_averages_by_decade$Average_Attempted_Threes/fga_averages_by_decade$Average_Attempted_Field_Goals*100)
percentage_of_fgas_tps
```
```{r}
#Scoring Average By Decade
averages_over_years_Points <- function(start,end) {
  year_string <- paste(as.character(start),as.character(end),sep = '-')
  df <-  averages %>% filter((Year >= start) & (Year <= end)) %>% group_by(Status) %>% summarize(Average_PTS = mean(PTS)) %>% mutate(Years = year_string) %>% select(Years,Status, Average_PTS)
  df 
}

i <-  0
pts_averages_by_decade = data.frame()
while (i < 3) {
  start <-  1980+10*i
  end <- 1989+10*i
  pts_averages_by_decade <- rbind(pts_averages_by_decade,averages_over_years_Points(start,end))
  i = i + 1
}
pts_averages_by_decade <- rbind(pts_averages_by_decade,averages_over_years_Points(2010,2017))
pts_averages_by_decade
```


```{r}
#Percentage of Point From Three Pointers
percentage_of_pts_tps_by_decade <- data.frame(Years = years,Status = status,Percentage_of_Points_TPs = 3*100*tp_averages_by_decade$Average_Made_Threes/pts_averages_by_decade$Average_PTS)
percentage_of_pts_tps_by_decade 
```

```{r}
#Two Point Field Goal Percentage By Decade
two_point_field_goals_made_by_decade <- fg_averages_by_decade$Average_Made_Field_Goals- tp_averages_by_decade$Average_Made_Threes
two_point_field_goals_attempted_by_decade <- fga_averages_by_decade$Average_Attempted_Field_Goals - tpa_averages_by_decade$Average_Attempted_Threes
twop_fgp_by_decade <- data.frame(Years = years, Status = status,Two_Point_FGP =  100*two_point_field_goals_made_by_decade/two_point_field_goals_attempted_by_decade) 
twop_fgp_by_decade

```




```{r}
#Free Throws Made By Decade

averages_over_years_FT <- function(start,end) {
  year_string <- paste(as.character(start),as.character(end),sep = '-')
  df <-  averages %>% filter((Year >= start) & (Year <= end)) %>% group_by(Status) %>% summarize(Average_FT = mean(FT)) %>% mutate(Years = year_string) %>% select(Years,Status, Average_FT)
  df 
}

i <-  0
ft_averages_by_decade = data.frame()
while (i < 3) {
  start <-  1980+10*i
  end <- 1989+10*i
  ft_averages_by_decade <- rbind(ft_averages_by_decade,averages_over_years_FT(start,end))
  i = i + 1
}
ft_averages_by_decade <- rbind(ft_averages_by_decade,averages_over_years_FT(2010,2017))

ft_averages_by_decade
```

```{r}
#Free Throws Attempts By Decade

averages_over_years_FTA <- function(start,end) {
  year_string <- paste(as.character(start),as.character(end),sep = '-')
  df <-  averages %>% filter((Year >= start) & (Year <= end)) %>% group_by(Status) %>% summarize(Average_FTA = mean(FTA)) %>% mutate(Years = year_string) %>% select(Years,Status, Average_FTA)
  df 
}

i <-  0
fta_averages_by_decade = data.frame()
while (i < 3) {
  start <-  1980+10*i
  end <- 1989+10*i
  fta_averages_by_decade <- rbind(fta_averages_by_decade,averages_over_years_FTA(start,end))
  i = i + 1
}
fta_averages_by_decade <- rbind(fta_averages_by_decade,averages_over_years_FTA(2010,2017))

fta_averages_by_decade

```

```{r}
#Free Throw Percentage By Decade
ftp_by_decade <- data.frame(Years = years, Status = status, FTP = ft_averages_by_decade$Average_FT/fta_averages_by_decade$Average_FTA*100)
ftp_by_decade
```
```{r}
#Scoring Breakdown 
points_breakdown <- data.frame(Years=years, Status = status, Points_Average = pts_averages_by_decade$Average_PTS,Percent_Points_Twos = 2*fg_averages_by_decade$Average_Made_Field_Goals/pts_averages_by_decade$Average_PTS*100,Percent_Points_Three = percentage_of_pts_tps_by_decade$Percentage_of_Points_TPs,Percent_Points_FTs = 100*ft_averages_by_decade$Average_FT/pts_averages_by_decade$Average_PTS)
points_breakdown
```


```{r}
write_csv(pts_averages_by_decade,'decade_averages/pts_averages_by_decade.csv')
write_csv(fg_averages_by_decade,'decade_averages/fg_averages_by_decade.csv')
write_csv(fga_averages_by_decade,'decade_averages/fg_attempted_by_decade.csv')
write_csv(fgp_by_decade,'decade_averages/fgp_by_decade.csv')
write_csv(tp_averages_by_decade,'decade_averages/tp_averages_by_decade.csv')
write_csv(tpa_averages_by_decade,'decade_averages/tpa_averages_by_decade.csv')
write_csv(tpp_by_decade,'decade_averages/tpp_by_decade.csv')
write_csv(percentage_of_fgs_tps,'decade_averages/percentage_of_fgs_tps_by_decade.csv')
write_csv(percentage_of_fgas_tps,'decade_averages/percentage_of_fgas_tps_by_decade.csv')
write_csv(percentage_of_pts_tps_by_decade ,'decade_averages/percentage_of_pts_tps_by_decade.csv')
write_csv(twop_fgp_by_decade,'decade_averages/twop_fgp_by_decade.csv')
write_csv(fta_averages_by_decade,'decade_averages/fta_averages_by_decade.csv')
write_csv(ft_averages_by_decade,'decade_averages/ft_averages_by_decade.csv')
write_csv(ftp_by_decade,'decade_averages/ftp_by_decade.csv')

```

```{r}
#Stats By Decade, Combined
stats_by_decade <- data.frame(Years = years, Status = status)
stats_by_decade <- cbind(stats_by_decade,select(pts_averages_by_decade,-c(Years,Status)),select(fg_averages_by_decade))
stats_by_decade <- cbind(stats_by_decade,select(fg_averages_by_decade,-c(Years,Status)))
stats_by_decade <- cbind(stats_by_decade,select(fga_averages_by_decade,-c(Years,Status)))
stats_by_decade <- cbind(stats_by_decade,select(fgp_by_decade,-c(Years,Status)))
stats_by_decade <- cbind(stats_by_decade,select(twop_fgp_by_decade,-c(Years,Status)))
stats_by_decade <- cbind(stats_by_decade,select(tp_averages_by_decade,-c(Years,Status)))
stats_by_decade <- cbind(stats_by_decade,select(tpa_averages_by_decade,-c(Years,Status)))
stats_by_decade <- cbind(stats_by_decade,select(tpp_by_decade,-c(Years,Status)))
stats_by_decade <- cbind(stats_by_decade,select(ft_averages_by_decade,-c(Years,Status)))
stats_by_decade <- cbind(stats_by_decade,select(fta_averages_by_decade,-c(Years,Status)))
stats_by_decade <- cbind(stats_by_decade,select(ftp_by_decade,-c(Years,Status)))
stats_by_decade <- cbind(stats_by_decade,select(percentage_of_fgs_tps,-c(Years,Status)))
stats_by_decade <- cbind(stats_by_decade,select(percentage_of_fgas_tps,-c(Years,Status)))
stats_by_decade <- cbind(stats_by_decade,select(points_breakdown,-c(Years,Status,Points_Average)))

stats_by_decade






```

```{r}
write_csv(stats_by_decade,'stats_by_decade.csv')
```
